---
title: 'RV466/EID005: Magpix Data Report'
author: "Nicole Dear"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# SET UP WORKSPACE -----

# load packages
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyr)
library(expss)
library(ggpubr)
library(stringr)
library(tidyverse)
library(table1)
library(printr)
library(flextable)
library(knitr)
# library(fmsb) # for radar (polar) charts
# theme_set(theme_pubr())

# set working directory
setwd("G:/DCAC/DCAC_EIDB/EID005_RV466/Analysis/MagPix")

# load magpix data (these come from Fauci, exported as excel)
magpix <- read_excel("G:/DCAC/DCAC_EIDB/EID005_RV466/Analysis/MagPix/RV466_magpix_clean.xlsx")

```


### Description
This report summarizes the RV466/EID005 Magpix (IgG) data.


### Methods
Descriptive methods were used to analyze Magpix IgG data, including frequencies and percentages for categorical variables. No statistical hypothesis testing was performed. A positive result was defined as a signal-to-noise ratio value >20.


```{r include=FALSE}

# DATA CLEANING -----

# create variable for population (healthy controls [n=8] vs acutely ill)
magpix$control <- ifelse(magpix$subj_id=="2019001"|magpix$subj_id=="2019002"|magpix$subj_id=="2019003"|
                           magpix$subj_id=="2019004"|magpix$subj_id=="2019005"|magpix$subj_id=="2019006"|
                           magpix$subj_id=="2019007"|magpix$subj_id=="2019008", 1, 0) 

# create country variable
magpix$country <- ifelse(grepl('10', substr(magpix$subj_id,1,3)),1,
                         ifelse(grepl('20', substr(magpix$subj_id,1,3)),2,NA))

# add labels
magpix$control <- factor(magpix$control,
                    levels = c(0,1),
                    labels = c("Acutely ill", "Healthy controls"))

magpix$country <- factor(magpix$country,
                    levels = c(1,2),
                    labels = c("Nigeria", "Ghana"))

# check frequencies
addmargins(table(magpix$country,magpix$control))

# create flags (per Keersten Ricks and lab team, a S/N of >20 indicates a 'positive' result)
magpix$RVFV_POS <- ifelse(magpix$RVFV>20, 1, 0)
magpix$PANA_POS <- ifelse(magpix$Panalphavirus>20, 1, 0)
magpix$PANF_POS <- ifelse(magpix$Panflavivirus>20, 1, 0)
magpix$MARV_POS <- ifelse(magpix$MARV>20, 1, 0)
magpix$EBOV_POS <- ifelse(magpix$EBOV>20, 1, 0)
magpix$LASV_POS <- ifelse(magpix$LASV>20, 1, 0)
magpix$CCHFV_POS <- ifelse(magpix$CCHFV>20, 1, 0)

# write.csv(magpix, file = "magpix_flags.csv")

```


### Table 1. Percentage of participants with signal-to-noise ratio >20

```{r echo=FALSE}

# TABLE 1 -----

magpix_table <- magpix

magpix_table$RVFV_POS <- factor(magpix_table$RVFV_POS,
                    levels = c(0,1),
                    labels = c("Negative", "Positive"))

magpix_table$PANA_POS <- factor(magpix_table$PANA_POS,
                    levels = c(0,1),
                    labels = c("Negative", "Positive"))

magpix_table$PANF_POS <- factor(magpix_table$PANF_POS,
                    levels = c(0,1),
                    labels = c("Negative", "Positive"))

magpix_table$MARV_POS <- factor(magpix_table$MARV_POS,
                    levels = c(0,1),
                    labels = c("Negative", "Positive"))

magpix_table$EBOV_POS <- factor(magpix_table$EBOV_POS,
                    levels = c(0,1),
                    labels = c("Negative", "Positive"))

magpix_table$LASV_POS <- factor(magpix_table$LASV_POS,
                    levels = c(0,1),
                    labels = c("Negative", "Positive"))

magpix_table$CCHFV_POS <- factor(magpix_table$CCHFV_POS,
                    levels = c(0,1),
                    labels = c("Negative", "Positive"))

label(magpix_table$RVFV_POS) <- "RVFV"
label(magpix_table$PANA_POS) <- "Alphavirus spp."
label(magpix_table$PANF_POS) <- "Flavivirus spp."
label(magpix_table$MARV_POS) <- "MARV"
label(magpix_table$EBOV_POS) <- "EBOV"
label(magpix_table$LASV_POS) <- "LASV"
label(magpix_table$CCHFV_POS) <- "CCHFV"

pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}


t1 <- table1(~ RVFV_POS + PANA_POS + PANF_POS + MARV_POS + EBOV_POS + LASV_POS + CCHFV_POS | 
               country, data = magpix_table, overall=c(right="Total"), extra.col=list(`P-value`=pvalue))

t1flex(t1)

# reference
# https://cran.r-project.org/web/packages/table1/vignettes/table1-examples.html

```


### Figure 1. Bar graph of pathogen positivity by country

```{r echo=FALSE, warning = FALSE}

# SUMMARIZE DATA IN BAR GRAPH -----

# restructure data
magpix_long <- gather(magpix, path, pos, RVFV_POS:CCHFV_POS, factor_key=TRUE)

magpix_long_mod <- magpix_long %>%
  mutate(pathogen = case_when(path=="RVFV_POS" ~ 1,
                              path=="PANA_POS" ~ 2,
                              path=="PANF_POS" ~ 3,
                              path=="MARV_POS" ~ 4,
                              path=="EBOV_POS" ~ 5,
                              path=="LASV_POS" ~ 6,
                              path=="CCHFV_POS" ~ 7))

magpix_long_mod$pathogen <- factor(magpix_long_mod$pathogen,
                                      levels = c(1,2,3,4,5,6,7),
                                      labels = c("RVFV", "ALPHA", "FLAVI",
                                                 "MARV", "EBOV", "LASV", "CCHFV"))

# calculate counts of those positive for each pathogen
pos <- magpix_long_mod %>%
  group_by(country, pathogen) %>%
  summarise(tot_pos = sum(pos), 
            n = n())

pos$percent <- round(100*(pos$tot_pos/pos$n),2)

# create labels
pos$label <- paste(pos$tot_pos, " (", round(pos$percent, digits = 1), "%)", sep = "")

# create color palette
#cpalette <- c("#375623","#ABA72B")
cpalette <- c("black","darkgrey")

# plot
bar <- ggplot(pos, aes(x = pathogen, y = percent, fill=country)) +
  geom_col(position="dodge") +
  geom_text(aes(label = label), size = 4, position = position_dodge(width = 1.1), vjust = -0.6) +
  scale_fill_manual(values = cpalette) +
  theme_pubclean() +
  xlab("Pathogen") +
  ylab("Percent") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        legend.title=element_text(size=16), 
        legend.text=element_text(size=14)) +
  guides(fill=guide_legend(title="Country")) +
  lims(y = c(0, 100))
  # + labs(caption = "Note: N=256 participants (n=154 in Nigeria and n=102 participants in Ghana) \n 
  #      underwent Magpix testing for each pathogen", size = 18)

png(file="RV466_Magpix_Fig1_gs.png", width=800, height=600)
bar
dev.off()

```


### Figure 2. Distribution of S/N values by pathogen and country (histogram)

```{r echo=FALSE, warning = FALSE}

# restructure S/N ratio data
magpix_sn_long <- gather(magpix, path, sn_value, RVFV:CCHFV, factor_key=TRUE)

magpix_sn_long_mod <- magpix_sn_long %>%
  mutate(pathogen = case_when(path=="RVFV" ~ 1,
                              path=="Panalphavirus" ~ 2,
                              path=="Panflavivirus" ~ 3,
                              path=="MARV" ~ 4,
                              path=="EBOV" ~ 5,
                              path=="LASV" ~ 6,
                              path=="CCHFV" ~ 7))

magpix_sn_long_mod$pathogen <- factor(magpix_sn_long_mod$pathogen,
                                levels = c(1,2,3,4,5,6,7),
                                labels = c("RVFV","Panalphavirus","Panflavivirus","MARV","EBOV","LASV","CCHFV"))

# check ranges of S/N values
t1 <- table1(~ RVFV + Panalphavirus + Panflavivirus + MARV + EBOV + LASV + CCHFV | 
               country, data = magpix_table, overall=c(right="Total"))

t1flex(t1)

# histogram of S/N ratio values by pathogen type and country
sn_hist <- gghistogram(magpix_sn_long_mod, x = "sn_value", binwidth = 8, 
                       # rug = TRUE,
                       color = "pathogen", fill = "pathogen",
                       xlab = "S/N Ratio",
                       ylab = "Count",
                       title = "Distribution of S/N Values by Pathogen and Country") +
  theme(legend.position="right", legend.box="vertical") +
  theme(legend.title = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_vline(aes(xintercept = 20), col='red', linewidth=0.5, linetype=3) +
  # annotate("text", x = 40, y = -10, label = "S/N>20", size=3) +  
  labs(caption = "Note: The red dashed line indicates a S/N ratio value of 20; binwidth is 8 (n=32 bins).", 
       size = 12) +
  coord_cartesian(ylim = c(0, 140), clip = "off") +
  facet_grid(country ~ .)

#png(file="RV466_Magpix_Fig2.png", width=700, height=700)
sn_hist
#dev.off()
```


### Figure 3. Patterns of Pathogen Positivity

```{r echo=FALSE}

# create 'heatmap' of different positivity patterns

magpix$RVFV_pos_c <- ifelse(magpix$RVFV>20, "RVFV", "")
magpix$Panalphavirus_pos_c <- ifelse(magpix$Panalphavirus>20, "Panalphavirus", "")
magpix$Panflavivirus_pos_c <- ifelse(magpix$Panflavivirus>20, "Panflavivirus", "")
magpix$MARV_pos_c <- ifelse(magpix$MARV>20, "MARV", "")
magpix$EBOV_pos_c <- ifelse(magpix$EBOV>20, "EBOV", "")
magpix$LASV_pos_c <- ifelse(magpix$LASV>20, "LASV", "")
magpix$CCHFV_pos_c <- ifelse(magpix$CCHFV>20, "CCHFV", "")

magpix$pattern <- paste(magpix$RVFV_pos_c, magpix$Panalphavirus_pos_c, magpix$Panflavivirus_pos_c,
                        magpix$MARV_pos_c, magpix$EBOV_pos_c, magpix$LASV_pos_c, magpix$CCHFV_pos_c, 
                        sep = " ", collapse = NULL)

# remove extra spaces between words
magpix$pattern <- trimws(magpix$pattern)
magpix$pattern <- gsub("\\s+"," ", magpix$pattern)

mp <- as.data.frame(sort(table(magpix$pattern), decreasing=T))
mp$tot <- 256
mp$percent <- round(100*(mp$Freq/mp$tot), 2)

mp$Var1[mp$Var1==""] <- NA
print(mp)

# write.csv(mp, file = "magpix_patterns.csv")


# tree map
library(treemap)

# create labels
mp$id <- 1:nrow(mp)

mp$label[mp$id==1] <- "Panflavivirus"
mp$label[mp$id==2] <- "None"
mp$label[mp$id==3] <- "Panalphavirus, Panflavivirus"
mp$label[mp$id==4] <- "RVFV, Panalphavirus, Panflavivirus"
mp$label[mp$id==5] <- "Panalphavirus"
mp$label[mp$id==6] <- "RVFV, Panflavivirus"
mp$label[mp$id==7] <- "LASV"
mp$label[mp$id==8] <- "Panalphavirus, Panflavivirus, EBOV"
mp$label[mp$id==9] <- "Panalphavirus, Panflavivirus, LASV"
mp$label[mp$id==10] <- "Panflavivirus, LASV"
mp$label[mp$id==11] <- "Panalphavirus, Panflavivirus, CCHFV"
mp$label[mp$id==12] <- "Panflavivirus, CCHFV"
mp$label[mp$id==13] <- "CCHFV"
mp$label[mp$id==14] <- "RVFV, Panflavivirus, LASV"
mp$label[mp$id==15] <- "Panalphavirus, LASV"
mp$label[mp$id==16] <- "Panflavivirus, EBOV"
mp$label[mp$id==17] <- "Panflavivirus, MARV"
mp$label[mp$id==18] <- "Panflavivirus, MARV, EBOV"
mp$label[mp$id==19] <- "RVFV, Panalphavirus, Panflavivirus, EBOV"
mp$label[mp$id==20] <- "RVFV, Panalphavirus, Panflavivirus, EBOV, CCHFV"
mp$label[mp$id==21] <- "RVFV, Panalphavirus, Panflavivirus, LASV"
mp$label[mp$id==22] <- "RVFV, Panflavivirus, EBOV"
mp$label[mp$id==23] <- "RVFV, Panflavivirus, EBOV, LASV, CCHFV"

# keep only patterns with at least 2 pathogens detected
mp2 <- mp[c(3:4,6,8:12,14:23),]

# keep top 10
top10 <- mp2[c(1:10),]

top10$label2 <- paste(top10$label, ": ", top10$Freq, " (", round(top10$percent, digits = 1), "%)", sep = "")

treemap(top10,
            index="label2",
            vSize="Freq",
            type="index",
            title=" ",
            fontsize.labels = 10)


# circle packing
library(packcircles)
library(viridis)

# Generate the layout. This function return a dataframe with one line per bubble. 
# It gives its center (x and y) and its radius, proportional of the value
packing <- circleProgressiveLayout(top10$Freq, sizetype='area')

# We can add these packing information to the initial data frame
top10 <- cbind(top10, packing)
 
# Check that radius is proportional to value. We don't want a linear relationship, since it is the AREA that must be proportionnal to the value
plot(top10$radius, top10$Freq)
 
# The next step is to go from one center + a radius to the coordinates of a circle that
# is drawn by a multitude of straight lines.
dat.gg <- circleLayoutVertices(packing, npoints=50)
 
# Make the plot
bubbles <- ggplot() + 
  geom_polygon(data = dat.gg, aes(x, y, group = id, fill=as.factor(id)), colour = "black", alpha = 0.6) +
  scale_fill_manual(values = magma(nrow(top10))) +
  geom_text(data = top10, aes(x, y, size=Freq, label = label)) +
  scale_size_continuous(range = c(1,4)) +
  theme_void() + 
  theme(legend.position="none") +
  coord_equal()

#png(file="RV466_Magpix_Fig3.png", width=800, height=600)
bubbles
#dev.off()

```


```{r echo=FALSE}

# DATA WRANGLING FOR MAPS -----
# DISTRICT LEVEL

# load data
mp_geo <- read_excel("G:/DCAC/DCAC_EIDB/EID005_RV466/Analysis/MagPix/magpix_flags.xlsx")

mp_geo_g <- mp_geo %>% subset(country=="Ghana")
mp_geo_n <- mp_geo %>% subset(country=="Nigeria")

# create summary measures for Ghana
mp_geo_g_summary <- mp_geo_g %>%
  group_by(ID) %>%
  summarise(RVFV_POS_TOT = sum(RVFV_POS),
            PANA_POS_TOT = sum(PANA_POS),
            PANF_POS_TOT = sum(PANF_POS),
            MARV_POS_TOT = sum(MARV_POS),
            EBOV_POS_TOT = sum(EBOV_POS),
            LASV_POS_TOT = sum(LASV_POS),
            CCHFV_POS_TOT = sum(CCHFV_POS),
            n = n())

mp_geo_g_summary$RVFV_POS_PROP <- mp_geo_g_summary$RVFV_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$PANA_POS_PROP <- mp_geo_g_summary$PANA_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$PANF_POS_PROP <- mp_geo_g_summary$PANF_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$MARV_POS_PROP <- mp_geo_g_summary$MARV_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$EBOV_POS_PROP <- mp_geo_g_summary$EBOV_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$LASV_POS_PROP <- mp_geo_g_summary$LASV_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$CCHFV_POS_PROP <- mp_geo_g_summary$CCHFV_POS_TOT/mp_geo_g_summary$n

write.csv(mp_geo_g_summary, file = "magpix_geo_ghana_summary.csv")

# create summary measures for Nigeria
mp_geo_n_summary <- mp_geo_n %>%
  group_by(ID) %>%
  summarise(RVFV_POS_TOT = sum(RVFV_POS),
            PANA_POS_TOT = sum(PANA_POS),
            PANF_POS_TOT = sum(PANF_POS),
            MARV_POS_TOT = sum(MARV_POS),
            EBOV_POS_TOT = sum(EBOV_POS),
            LASV_POS_TOT = sum(LASV_POS),
            CCHFV_POS_TOT = sum(CCHFV_POS),
            n = n())

mp_geo_n_summary$RVFV_POS_PROP <- mp_geo_n_summary$RVFV_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$PANA_POS_PROP <- mp_geo_n_summary$PANA_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$PANF_POS_PROP <- mp_geo_n_summary$PANF_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$MARV_POS_PROP <- mp_geo_n_summary$MARV_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$EBOV_POS_PROP <- mp_geo_n_summary$EBOV_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$LASV_POS_PROP <- mp_geo_n_summary$LASV_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$CCHFV_POS_PROP <- mp_geo_n_summary$CCHFV_POS_TOT/mp_geo_n_summary$n

write.csv(mp_geo_n_summary, file = "magpix_geo_nigeria_summary.csv")
```


```{r echo=FALSE}

# DATA WRANGLING FOR MAPS -----
# REGIONAL LEVEL

# load data
mp_geo <- read_excel("G:/DCAC/DCAC_EIDB/EID005_RV466/Analysis/MagPix/magpix_flags.xlsx")

# manual corrections of region names
mp_geo$region[mp_geo$region=="BENUE"] <- "Benue"
mp_geo$region[mp_geo$region=="Lagos state"] <- "Lagos"
mp_geo$region[mp_geo$region=="Lagos State"] <- "Lagos"
mp_geo$region[mp_geo$region=="Nigeria"] <- "Niger"

# split into two separate data sets, one for each country
mp_geo_g <- mp_geo %>% subset(country=="Ghana")
mp_geo_n <- mp_geo %>% subset(country=="Nigeria")

# create summary measures for Ghana
mp_geo_g_summary <- mp_geo_g %>%
  group_by(region) %>%
  summarise(RVFV_POS_TOT = sum(RVFV_POS),
            PANA_POS_TOT = sum(PANA_POS),
            PANF_POS_TOT = sum(PANF_POS),
            MARV_POS_TOT = sum(MARV_POS),
            EBOV_POS_TOT = sum(EBOV_POS),
            LASV_POS_TOT = sum(LASV_POS),
            CCHFV_POS_TOT = sum(CCHFV_POS),
            n = n())

mp_geo_g_summary$RVFV_POS_PROP <- mp_geo_g_summary$RVFV_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$PANA_POS_PROP <- mp_geo_g_summary$PANA_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$PANF_POS_PROP <- mp_geo_g_summary$PANF_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$MARV_POS_PROP <- mp_geo_g_summary$MARV_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$EBOV_POS_PROP <- mp_geo_g_summary$EBOV_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$LASV_POS_PROP <- mp_geo_g_summary$LASV_POS_TOT/mp_geo_g_summary$n
mp_geo_g_summary$CCHFV_POS_PROP <- mp_geo_g_summary$CCHFV_POS_TOT/mp_geo_g_summary$n

write.csv(mp_geo_g_summary, file = "magpix_geo_ghana_regional_summary.csv")

# create summary measures for Nigeria
mp_geo_n_summary <- mp_geo_n %>%
  group_by(region) %>%
  summarise(RVFV_POS_TOT = sum(RVFV_POS),
            PANA_POS_TOT = sum(PANA_POS),
            PANF_POS_TOT = sum(PANF_POS),
            MARV_POS_TOT = sum(MARV_POS),
            EBOV_POS_TOT = sum(EBOV_POS),
            LASV_POS_TOT = sum(LASV_POS),
            CCHFV_POS_TOT = sum(CCHFV_POS),
            n = n())

mp_geo_n_summary$RVFV_POS_PROP <- mp_geo_n_summary$RVFV_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$PANA_POS_PROP <- mp_geo_n_summary$PANA_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$PANF_POS_PROP <- mp_geo_n_summary$PANF_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$MARV_POS_PROP <- mp_geo_n_summary$MARV_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$EBOV_POS_PROP <- mp_geo_n_summary$EBOV_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$LASV_POS_PROP <- mp_geo_n_summary$LASV_POS_TOT/mp_geo_n_summary$n
mp_geo_n_summary$CCHFV_POS_PROP <- mp_geo_n_summary$CCHFV_POS_TOT/mp_geo_n_summary$n

write.csv(mp_geo_n_summary, file = "magpix_geo_nigeria_regional_summary.csv")

```
